### 2023-01-23 manuscript.Rmd
fixed some language issues

Based on four basic principles of a more responsible research assessment in academic hiring and promotion processes [@schonbrodt2022responsible], @gartner2022responsible suggest a specific implementation to broaden the range of what constitutes research contributions. Gaertner et al. provide evaluation schemes for published manuscripts, published datasets, and software contributions.
Why are research software contributions important academic contributions? We would like to respond with a quote from the Science Code Manifesto (http://sciencecodemanifesto.org/):
"Software is an essential research product, and the effort to produce, maintain, adapt, and 
curate code must be recognized."
and
"Software is a cornerstone of science. Without software, twenty-first century science would be impossible."
Yet, despite the heavy reliance on computational infrastructure, the current academic infrastructure does not adequately incentivice good software engineering practice that we would expect from a research software engineer [@baxter2012research]. 
 Here, we specifically comment on the proposed evaluation scheme for research software contributions (Table 3 of @gartner2022responsible). We comment on the dimensions metrics, lines of code, software engineering, impact, peer-review, usability.

@@

Date of first fully functional public version is a central field in the proposal as it serves as a denominator in computing citations relative to software age. In our own experience, research software often evolves very differently to the standards that software engineering prescribes. In best-practice software development, the first step is software specification or requirements engineering. This is the process of understanding and defining what features and services are required from the product and how they are to be offered. This involves research on existing solutions, discussions with users, user interface testing with mockups or user observations while testing prototypes. A requirement specification documents the expected functionality, such that development time and cost can be estimated and requirements can be formally checked with the final product for consistency and completeness. From our own experience, hardly any of these steps apply to scientific software development. Scientific software is often developed as a side-project fuelled by enthusiasm instead of proper funding. Some pieces of software start as testbed implementation of a newly proposed methodology, others come to life as wrappers or graphical interfaces for other packages, some others were never meant to be shared but are so useful that they became a package on their own.

@@

Complementing the quantitative aspect of impact, we suggest to introduce a mandatory qualitative judgement on the impact of the software, which fits with what the others proposed as indicator "merit / impact statement". In our view, researchers should be requested to indicate at least one (and up to three) empirical research projects that directly benefitted from the software contributions they made. We believe this is a fair assessment of the real impact no matter whether someone contributed documentation/tutorial to run certain types of models, whether someone contributed a specific feature that enabled a given research project to run their analysis, or whether someone is the author of the entire package.

@@

We strongly oppose the lines of code (LOC) metric that was proposed to measure effort. Lines of code is a common metric in software development useful to gauge the size of an entire code base or its change over time. There are several problems in using this indicator for effort. First of all, lines of code is an ambiguous metric. It can be calculated differently depending on whether comments do count or whether physical lines of code or logical lines of code are counted.  Second, lines of codes for code contributions highly depends on the programming language, mastery and personal programming style. In particular, large LOC of a researcher may simply mean that they write unefficient and repetitive code, one of the great sins of programming. A feature of good software is modularity because it enable reuse of functions both inside and outside the project. A great example is the CRAN library that enables packages to built upon other packages and reuse their functionality. However, if LOC were a metric that a researcher are compared with, it encourages researchers to copy&paste chunks of code from other packages instead of importing them in their project.

@@

Similar to peer-review for manuscripts, software code is amenable to peer-review. It is common
practice in corporate software development, to have peers review code. For example, if a code base is changed, it is good practice to request a change with the developer team before actually changing the code base. The change is reviewed by experts and/or maintainers before it is integrated in the active branch of development. In developing these guidelines, we see the opportunity to foster such good practices and reward researchers that actively engage in reviewing computer code. Software peer-review is increasingly becoming recognized as a means to enable better research software. Peer-review can happen in the context of dedicated software journal (e.g, Journal of Statistical Software), as part of the standard review process (for guidelines, see https://www.carlboettiger.info/2013/06/13/what-I-look-for-in-software-papers.html), or as part of submissions to software archives that guarantuee certain standards (e.g., rOpenSci). We suggest that active participation in the review of scientific software, independent of the involvement or maintance of specific software products, should be a dimension of positive evaluation.

@@

In sum, we are thankful to Gaertner et al. for highlighting the importance of research software contributions as academically valuable products. Specifically, we value ... However, we are critical of ... for the reasons mentioned here. We propose to ...

>>>>

Based on four basic principles of a more responsible research assessment in academic hiring and promotion processes [@schonbrodt2022responsible], @gartner2022responsible suggested a specific implementation of these principles to broaden the range of what constitutes research contributions. Gaertner et al. provide evaluation schemes for published manuscripts, published datasets, and software contributions. 
We specifically endorse the increased emphasis on research software as recognizable and laudable scientific contribution.
Why are research software contributions important scientific contributions? We would like to respond with a quote from the Science Code Manifesto (http://sciencecodemanifesto.org/):
"Software is an essential research product, and the effort to produce, maintain, adapt, and 
curate code must be recognized."
and
"Software is a cornerstone of science. Without software, twenty-first century science would be impossible."
Yet, despite the heavy reliance on computational infrastructure, the current academic infrastructure does not adequately incentivise software development and, specifically, good software engineering practice that we would expect from a research software engineer [@baxter2012research]. 
 Here, we specifically comment on the proposed evaluation scheme for research software contributions (Table 3 of @gartner2022responsible). We comment on the dimensions metrics, lines of code, software engineering, impact, peer-review, usability.

@@

<!-- This needs to be shortened a lot-->
Date of first fully functional public version is a central field in the proposal as it serves as a denominator in computing citations relative to software age. In our own experience, research software often evolves very differently to the standards that software engineering prescribes. In best-practice software development, the first step is software specification or requirements engineering. This is the process of understanding and defining what features and services are required from the product and how they are to be offered. This involves research on existing solutions, discussions with users, user interface testing with mockups or user observations while testing prototypes. A requirement specification documents the expected functionality, such that development time and cost can be estimated and requirements can be formally checked with the final product for consistency and completeness. From our own experience, hardly any of these steps apply to scientific software development. Scientific software is often developed as a side-project fuelled by enthusiasm instead of proper funding. Some pieces of software start as testbed implementation of a newly proposed methodology, others come to life as wrappers or graphical interfaces for other packages, some others were never meant to be shared but are so useful that they became a package on their own.

@@

Complementing the quantitative aspect of impact, we suggest to introduce a mandatory qualitative judgement on the impact of the software, which fits with what @gartner2022responsible proposed as indicator "merit / impact statement". In our view, researchers should be requested to indicate at least one (and up to three) empirical research projects that directly benefited from the software contributions they made. We believe this is a fair assessment of the real impact no matter whether someone contributed documentation/tutorial to run certain types of models, whether someone contributed a specific feature that enabled a given research project to run their analysis, or whether someone is the author of the entire package.

@@

We discourage the lines of code (LOC) metric that was proposed to measure effort. Lines of code is a common metric in software development useful to gauge the size of an entire code base or its change over time. There are several problems in using this indicator for effort. First of all, lines of code is an ambiguous metric. It can be calculated differently depending on whether comments do count or whether physical lines of code or logical lines of code are counted.  Second, lines of codes for code contributions highly depends on the programming language, mastery and personal programming style. In particular, large LOC of a researcher may simply mean that they write inefficient and repetitive code, one of the great sins of programming. A feature of good software is modularity because it enable reuse of functions both inside and outside the project. A great example is the CRAN library that enables packages to built upon other packages and reuse their functionality. However, if LOC were a metric that a researcher are compared with, it encourages researchers to copy&paste chunks of code from other packages instead of importing them in their project.

@@

Similar to peer-review for manuscripts, software code is amenable to peer-review. It is common
practice in corporate software development, to have peers review code. For example, if a code base is changed, it is good practice to request a change with the developer team before actually changing the code base. The change is reviewed by experts and/or maintainers before it is integrated in the active branch of development. In developing these guidelines, we see the opportunity to foster such good practices and reward researchers that actively engage in reviewing computer code. Software peer-review is increasingly becoming recognized as a means to enable better research software. Peer-review can happen in the context of dedicated software journal (e.g, Journal of Statistical Software), as part of the standard review process (for guidelines, see https://www.carlboettiger.info/2013/06/13/what-I-look-for-in-software-papers.html), or as part of submissions to software archives that guarantee certain standards (e.g., rOpenSci). We suggest that active participation in the review of scientific software, independent of the involvement or maintenance of specific software products, should be a dimension of positive evaluation.

@@

In sum, we are thankful to  @gartner2022responsible for highlighting the importance of research software contributions as scientifically valuable products. Specifically, we value the distinction of different classes of software projects (e.g., small scripts vs large-scale team efforts), ....  However, we are critical of the proposed quantitative impact measures. For the reasons mentioned here, We propose to emphasize that qualitative statement on impact, in which software authors should argue in how far their project paved the way for other scientific endeavours.
